---
title: "R Notebook"
output: html_notebook
---


```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
library(wesanderson)
library(ggpubr)
library(ggsci)
pal <- wes_palette("Zissou1", 100, type = "continuous")

source("../run/tools.R")
```


```{r}
output_path <- "../data/output/classifier_results_by_feature_group_predictions/" 
data <- read.csv(paste0(output_path, "result_pred_100000.csv"))
data_0 <- read.csv(paste0(output_path, "result_pred_100000_0.csv"))
data <- rbind(data, data_0)
rm(data_0)
```



### Mean correlation between predictions

```{r}
# Calculations

get_corr_table <- function(df, y){
  df1 <- df %>%
    select(X1, features_group, rn) %>%
    dcast(rn ~ features_group, value.var = "X1") %>%
    select(-rn) %>%
    cor(method="spearman") %>%
    melt() %>%
    mutate(
      spearman_corr = if_else(is.na(value), 0, value)
    ) %>%
    select(-value)
  return(df1)
}

grouped_corr <- data %>%
  group_by(cancer_type, agg_level, win_len, repeats) %>%
  group_modify(get_corr_table)
```

Mean in 30 repeats

```{r}
# write.csv(grouped_corr, "../data/adhoc/multipred_corr.csv")
grouped_corr <- read.csv("../data/adhoc/multipred_corr.csv")

mean_corr <- grouped_corr %>%
  group_by(cancer_type, agg_level, win_len, Var1, Var2) %>%
  summarize(
    mean_spearman_corr = mean(spearman_corr),
    median_spearman_corr = median(spearman_corr)
    ) %>%
  inner_join(
    proper_feature_group_names %>% select(feature_group_name, feature_group_name_proper), by=c("Var1"="feature_group_name")
  ) %>%
  rename("Var1_fg"="feature_group_name_proper") %>%
  inner_join(
    proper_feature_group_names %>% select(feature_group_name, feature_group_name_proper), by=c("Var2"="feature_group_name")
  ) %>%
  rename("Var2_fg"="feature_group_name_proper")

mean_corr$Var1_fg <- factor(mean_corr$Var1_fg, levels = c("TAD", "TF", "DNase", "HM", "methyl", "non-B DNA","region")) 
mean_corr$Var2_fg <- factor(mean_corr$Var2_fg, levels = c("TAD", "TF", "DNase", "HM", "methyl", "non-B DNA","region")) 

g1 <- ggplot(mean_corr %>% filter(agg_level == "99.") %>% filter(Var1_fg != Var2_fg), aes(x=Var1_fg, y=Var2_fg, fill=mean_spearman_corr))+
  geom_tile() + 
  facet_wrap(~cancer_type, nrow=2) +
  scale_fill_gradientn(colours = pal)+
  xlab("Feature group") + 
  ylab("Feature group") +
  theme_light()+
  labs(fill = "Spearman correlation")+
  theme(
    legend.position = "bottom",
    legend.title = element_text(hjust = 0, size = 10),
    axis.text.x = element_text(angle = 40, hjust = 1, size=8),
    axis.text.y = element_text(size=8)
    )+
  scale_x_discrete(limits=c("TAD", "DNase", "TF",  "HM", "methyl", "non-B DNA","region")) +
  scale_y_discrete(limits=c("region",  "non-B DNA", "methyl" , "HM", "TF","DNase", "TAD"))

g1
```


Additionally it is of interest here to investigate whether the different feature group models identify the same or different breakpoints hotspots regions. Firstly, we calculated mean Spearman correlation between predictions for each pair of feature groups (Fig.). 

```{r}
mean_corr$gr <- paste(mean_corr$Var1_fg, mean_corr$Var2_fg, sep="_")
mean_corr %>% 
  filter(Var1_fg != Var2_fg) %>%
  group_by(gr) %>%
  summarize(mc = mean(mean_spearman_corr)) %>%
  arrange(desc(mc))
```


The highest mean over all cancer types correlation is observed for histones modifications based models predictions that are positively associated with TF and methylation based models predictions (with correlation of 0.47 and 0.46 respectively). 
On opposite, TAD-based models generate the most specific predictions with maximal correlation of 0.045 with TF-based predictions.


```{r}
mean_corr %>% filter(Var1_fg != Var2_fg) %>%
  group_by(cancer_type, gr) %>%
  summarize(mc = mean(mean_spearman_corr)) %>%
  mutate(r = dense_rank(desc(mc))) %>%
  filter(r == 1) %>%
  arrange(r, desc(mc))
```

### Intersection

```{r}
get_recall <- function(test_pred, y){
  n <- round(nrow(test_pred) * 0.05)
  df1 <- test_pred %>%
    select(X1, rn) %>%
    arrange(desc(X1)) %>%
    mutate(
      label = if_else(row_number() <= n, 1, 0)
    ) %>%
    filter(
      label == 1
    ) %>%
    select(rn, label)

  return(df1)
}

grouped_pred <- data %>%
  group_by(cancer_type, agg_level, win_len, repeats, features_group) %>%
  group_modify(get_recall)
# write.csv(grouped_pred, "../data/adhoc/multipred_pred.csv")

all_targets <- data %>%
  filter(target == "X1") %>%
  select(cancer_type, agg_level, win_len, repeats, rn, target) %>%
  unique()
all_targets$target <- 1
write.csv(all_targets, "../data/adhoc/multipred_targets.csv")

# rm(data)
```

```{r}
grouped_pred_0 <- read.csv("../data/adhoc/multipred_pred.csv")
all_targets <- read.csv("../data/adhoc/multipred_targets.csv")

gr <- unique(grouped_pred_0$features_group)
grouped_pred <- grouped_pred_0 %>%
  dcast(cancer_type + agg_level + win_len + repeats + rn ~ features_group, value.var = "label")

# drop tad predictions
grouped_pred$tad <- NULL
gr <- setdiff(gr, c("tad"))

grouped_pred <- grouped_pred %>%
  left_join(all_targets, by=c("cancer_type", "agg_level", "win_len", "repeats", "rn"))
grouped_pred[is.na(grouped_pred)] <- 0

grouped_pred$total_pred <- rowSums(grouped_pred[as.character(gr)])
grouped_pred

grouped_pred_summary <- grouped_pred %>%
  filter(total_pred > 0) %>%
  group_by(cancer_type, agg_level, win_len, repeats, total_pred) %>%
  summarize(
    n_preds = n(),
    n_hsp = sum(target)
    ) %>%
  mutate(
    ratio = n_hsp / n_preds
  )

```


### Effect of combined predictiona

```{r}
grouped_pred_summary %>%
 group_by(total_pred) %>%
  summarize(mean(n_preds))
```


```{r}
grouped_pred_summary_plot <- grouped_pred_summary %>%
  filter(total_pred < 4)
grouped_pred_summary_plot$total_pred <- as.character(grouped_pred_summary_plot$total_pred)
grouped_pred_summary_plot <- grouped_pred_summary_plot[grouped_pred_summary_plot$agg_level == "99.", ]

mean_preds = grouped_pred_summary_plot %>%
  group_by(cancer_type, total_pred) %>%
  summarize(
    mean_n_preds = round(mean(n_preds), 0),
    m_r = max(ratio)
    )
mean_preds
```

```{r}
grouped_pred_single <- grouped_pred %>%
  filter(total_pred == 1) %>%
  select(-c("total_pred"))

grouped_pred_single <- melt(
  data=grouped_pred_single, 
  id.vars=c("cancer_type", "agg_level", "win_len", "repeats", "rn","target")
  ) %>%
  rename("feature_group"="variable") %>%
  filter(value==1) %>%
  group_by(cancer_type, agg_level, win_len, repeats, feature_group) %>%
  summarize(
    n_preds = sum(value),
    n_hsp = sum(target)
    ) %>%
  mutate(
    ratio = n_hsp / n_preds
  ) 

grouped_pred_single_st <- grouped_pred_single  %>%
  group_by(cancer_type, agg_level, win_len, feature_group) %>%
  summarize(
    ratio = mean(ratio),
    mean_n_preds = round(mean(n_preds), 0)
  ) %>%
  group_by(cancer_type, agg_level, win_len) %>%
  mutate(rank_ratio = dense_rank(desc(ratio))) %>%
  filter(rank_ratio == min(rank_ratio)) %>%
  select(-c("rank_ratio", "ratio"))

grouped_pred_single_best <- grouped_pred_single %>%
  inner_join(grouped_pred_single_st %>% select(-c("mean_n_preds")), by=c("cancer_type", "win_len", "feature_group", "agg_level")) %>%
  mutate(total_pred = "1")

grouped_pred_single_st <- grouped_pred_single_st %>%
  inner_join(
    grouped_pred_single_best %>%
      group_by(cancer_type, agg_level, win_len, feature_group) %>%
      summarize(m_r = max(ratio)) %>%
      ungroup() %>%
      select(-c("feature_group"))
  )
grouped_pred_single_st
```

```{r}
one_group_ratio <- grouped_pred_0 %>%
  left_join(all_targets %>% select(-X), by=c("cancer_type", "agg_level", "win_len", "repeats", "rn")) %>%
  mutate(
    target = if_else(is.na(target), 0, 1)
  ) %>%
  group_by(cancer_type, agg_level, win_len, repeats, features_group) %>%
  summarize(
    n_preds = sum(label),
    n_hsp = sum(target)
    ) %>%
  mutate(
    ratio = n_hsp / n_preds
  ) 

one_group_ratio_st <- one_group_ratio %>%
  group_by(cancer_type, agg_level, win_len, features_group) %>%
  summarize(
    ratio = mean(ratio)
  ) %>%
  filter(ratio == max(ratio))

one_group_ratio_best <- 
  one_group_ratio %>%
  inner_join(
    one_group_ratio_st %>% select(-ratio)
    ) %>%
  mutate(total_pred = "1")

```


```{r}
tot_pl <- rbind(
  grouped_pred_summary_plot %>% filter(total_pred != "1", agg_level == "99."),
  one_group_ratio_best %>% filter(agg_level == "99.")
)

tot_annot <- rbind.data.frame(
  one_group_ratio_best %>% 
    group_by(agg_level, win_len, cancer_type) %>%
    summarize(m_r = max(ratio)) %>%
    filter(agg_level == "99.") %>%
    ungroup() %>%
    select(-c("win_len", "agg_level")) %>%
    mutate(
      total_pred = "1",
      mean_n_preds=439
      ),
  mean_preds %>% filter(total_pred != "1")
)

g2 <- ggplot(tot_pl, 
  aes(x=total_pred, fill = total_pred,  y=ratio)) + 
  geom_boxplot() +
  theme_light()+
  facet_wrap(~cancer_type, scales="free", nrow=2)+
  xlab("Number of feature groups") + 
  ylab("Ratio of hotspots") +
  labs(fill = "Number of feature groups")+
  geom_text(data = tot_annot, aes(x = total_pred,  y=m_r + 0.01, label = mean_n_preds), size=3)+
  scale_fill_jco()+
  theme(
      legend.position = "bottom",
      legend.title = element_text(hjust = 0, size = 10),
  )

g2

g_f <- ggarrange(g1, g2,
          labels = c("A", "B"),
          ncol = 1, nrow = 2, heights = c(1, 1),
          font.label = list(size = 10, color = "black", face ="bold"))
g_f

# ggsave(filename = "Fig15_supp.tiff", plot = g_f, device = "tiff", path = "../paper_materials/plots_after_review/", dpi = 600, units = "in", width = 6, height = 5, scale = 2, compression = "lzw")
```

