---
title: "R Notebook"
output: html_notebook
---

## Boruta Feature Selection

Load data

```{r echo=F}
library(dplyr)
library(reshape2)

library(ggplot2)
library(wesanderson)
library("ggsci")

library(openxlsx)

library(formattable)

customYellow = 'yellow'
customBlue = '#1E90FF'
pal <- wes_palette("Zissou1", 100, type = "continuous")

source('../run/tools.R')
source('../run/features.R')

df <- read.csv("../data/output/boruta_res_100000.csv", row.names = 1, stringsAsFactors = F)
df <- unique(df)
df %>%
  group_by(agg_level) %>%
  summarize(n_distinct(cancer_type))
```

#### Analysis of features selection for task of hotspots prediction 

For each cancer type and labeling type there were 30 train/test splits which includes 10 different procedures for 5-iteration Boruta FS.

How many iterations were used in datasets?

```{r}
df_agg <- df %>%
  filter(feature_type == "real") %>%
  group_by(cancer_type, repeat., resample, iter) %>%
  summarize(
    n_selected = sum(hit),
    n_total = n()
    ) %>%
  mutate(
    ratio_selected = n_selected / n_total
  )

df_1_sample <- df_agg %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(max(iter) == 1) %>%
  select(cancer_type, repeat., resample) %>%
  unique()

df_11 <- df_agg %>%
  inner_join(df_1_sample, by = c("cancer_type", "repeat.", "resample")) %>%
  group_by(iter) %>%
  summarize(
    mean_n_selected = mean(n_selected),
    mean_ratio_selected = mean(ratio_selected)
  )

df_2_sample <- df_agg %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(max(iter) == 2) %>%
  select(cancer_type, repeat., resample) %>%
  unique()

df_21 <- df_agg %>%
  inner_join(df_2_sample, by = c("cancer_type", "repeat.", "resample")) %>%
  group_by(iter) %>%
  summarize(
    mean_n_selected = mean(n_selected),
    mean_ratio_selected = mean(ratio_selected)
  )

df_3_sample <- df_agg %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(max(iter) == 3) %>%
  select(cancer_type, repeat., resample) %>%
  unique()

df_31 <- df_agg %>%
  inner_join(df_3_sample, by = c("cancer_type", "repeat.", "resample")) %>%
  group_by(iter) %>%
  summarize(
    mean_n_selected = mean(n_selected),
    mean_ratio_selected = mean(ratio_selected)
  )

df_4_sample <- df_agg %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(max(iter) == 4) %>%
  select(cancer_type, repeat., resample) %>%
  unique()

df_41 <- df_agg %>%
  inner_join(df_4_sample, by = c("cancer_type", "repeat.", "resample")) %>%
  group_by(iter) %>%
  summarize(
    mean_n_selected = mean(n_selected),
    mean_ratio_selected = mean(ratio_selected)
  )

df_5_sample <- df_agg %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(max(iter) == 5) %>%
  select(cancer_type, repeat., resample) %>%
  unique()

df_51 <- df_agg %>%
  inner_join(df_5_sample, by = c("cancer_type", "repeat.", "resample")) %>%
  group_by(iter) %>%
  summarize(
    mean_n_selected = mean(n_selected),
    mean_ratio_selected = mean(ratio_selected)
  )

df_11$max_iter <- 1
df_21$max_iter <- 2
df_31$max_iter <- 3
df_41$max_iter <- 4
df_51$max_iter <- 5

df_stats <- rbind(
  df_11,
  df_21,
  df_31,
  df_41,
  df_51
  )

data.frame(
  n_iters = seq(1, 5),
  n_datasets = c(
    nrow(df_1_sample),
    nrow(df_2_sample), 
    nrow(df_3_sample), 
    nrow(df_4_sample),
    nrow(df_5_sample))
)

```

How namy features selected on each iteration?

```{r}
df_stats$max_iter <- as.character(df_stats$max_iter)
df_stats$iter <- as.character(df_stats$iter)

ggplot(df_stats, aes(x=max_iter, fill=iter, y=mean_n_selected)) + 
  geom_bar(stat="identity", color="black", position = "dodge")  + 
  scale_fill_jco()


```

Typically, 2-3 features are selected if there were less than 5 iterations (or when relatively big number of features were removed at the beginning of the procedure). In contrast, when all 5 iterations were used, lower percent of features were removed ar each iteration and in average 9 features were finally selected.

How namy features selected on each iteration in dependence on cancer type

```{r}
df_last_iter <- df_agg %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(iter == max(iter))

ggplot(df_last_iter, 
       aes(x=reorder(cancer_type, n_selected, FUN = median), fill=cancer_type, y=n_selected)) + 
  geom_violin() + 
  scale_fill_jco()
```

```{r}
df_last_iter %>%
  group_by(cancer_type) %>%
  summarize(median_iter = median(iter)) %>%
  arrange(median_iter)
```

```{r}
df_last_iter %>%
  group_by(cancer_type) %>%
  summarize(median_selected = median(n_selected)) %>%
  arrange(median_selected)
```

```{r}
df_last_iter %>%
  group_by(cancer_type) %>%
  filter(n_selected == 0) %>%
  summarize(n_zero_features = n()) %>%
  arrange(n_zero_features)
```

Two groups could be defined:

 1. Brain, liver, pancreatic, blood, skin - Median number of iterations is 1-4 with median number of selected features of 3-4. Besides, in 29-50 cases (from 300) no features were selected. All except skin cancer has labeling type 99.9%.
 2. Bone, breast, ovary, prostate, uterus - Median number of iterations is 5 with median number of selected features of 7-9. Besides, in 1-19 cases (from 300) no features were selected. Labeling types - 99% and 99.5%
 
```{r}
df %>%
  select(cancer_type, agg_level) %>%
  unique()
```


For each feature how many times it appeared in final feature set by cancer type

```{r, fig.width=10, fig.height=12}
df_all_hits <- df %>% 
  filter(feature_type == "real") %>%
  filter(hit == 1)

df_hits_last_iter <- df_all_hits %>%
  group_by(cancer_type, repeat., resample) %>%
  filter(iter == max(iter))

df_hits_last_iter$feature_n <- gsub(x=df_hits_last_iter$feature, pattern = "upper_", replacement = "")

feature_cols <- unique(df_hits_last_iter$feature)
df_feat_proper <- get_feature_df(feature_cols)

df_hits_last_iter <- df_hits_last_iter %>%
  inner_join(df_feat_proper)

df_hits_last_iter_agg <- df_hits_last_iter %>%
  group_by(cancer_type, feature, feature_group) %>%
  summarize(n_best = n())

f <- df_hits_last_iter_agg %>% 
  ungroup() %>% 
  select(feature, feature_group) %>%
  unique() %>%
  arrange(feature_group, feature)

df_hits_last_iter_agg$feature <- with(df_hits_last_iter_agg, factor(feature, levels = f$feature))

ggplot(df_hits_last_iter_agg, aes(x=cancer_type, y=feature,fill=n_best)) + 
  geom_tile() + 
  scale_fill_gradientn(colours = pal)
```

```{r}
df_hits_last_iter_agg$feature_new <- as.character(df_hits_last_iter_agg$feature)
df_hits_last_iter_agg$is_upper <- 0
df_hits_last_iter_agg$is_upper[grep(x = df_hits_last_iter_agg$feature_new, "upper_", value=F)] <- 1
df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "tf"]  <- gsub(x = df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "tf"], pattern = "cancer_liver_", "")
df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "tf"] <- gsub(x = df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "tf"], pattern = ".human", "")
df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "chromatin"] <- "DNase"
df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "methyl"] <- "methyl"
df_hits_last_iter_agg$feature_new[(df_hits_last_iter_agg$feature_group == "histones") & (df_hits_last_iter_agg$is_upper == 0)] <- unlist(lapply(df_hits_last_iter_agg$feature_new[(df_hits_last_iter_agg$feature_group == "histones") & (df_hits_last_iter_agg$is_upper == 0)], FUN = function(x) strsplit(x, split = "_")[[1]][[3]]))
df_hits_last_iter_agg$feature_new[(df_hits_last_iter_agg$feature_group == "histones") & (df_hits_last_iter_agg$is_upper == 1)] <- paste0("upper_", unlist(lapply(df_hits_last_iter_agg$feature_new[(df_hits_last_iter_agg$feature_group == "histones") & (df_hits_last_iter_agg$is_upper == 1)], FUN = function(x) strsplit(x, split = "_")[[1]][[4]])))
df_hits_last_iter_agg$feature_new[(df_hits_last_iter_agg$feature_group == "histones")] <- gsub(x = df_hits_last_iter_agg$feature_new[df_hits_last_iter_agg$feature_group == "histones"], pattern = ".human", "")

df_hits_last_iter_agg <- df_hits_last_iter_agg %>%
  inner_join(proper_feature_group_names, by=c("feature_group"="feature_group_name"))

df_hits_last_iter_agg <- df_hits_last_iter_agg %>%
  arrange(cancer_type, feature_group_name_proper, desc(feature_new))
df_hits_last_iter_agg
```

```{r, fig.width=10, fig.height=12}
df_hits_last_iter_agg1 <- df_hits_last_iter_agg %>%
  inner_join(
    df_hits_last_iter_agg %>%
    ungroup() %>%
    select(feature_group_name_proper, feature_new) %>%
    unique() %>%
    arrange(feature_group_name_proper, feature_new) %>%
    mutate(r = row_number())) %>%
  arrange(cancer_type, r)

g1 <- ggplot(df_hits_last_iter_agg1, aes(x=cancer_type, y=reorder(feature_new, -r), fill=n_best)) + 
  geom_tile() + 
  scale_fill_gradientn(colours = pal)+
  theme_light_custom(scale_fill = F)+
  xlab("Cancer type")+
  ylab("Feature")+
  labs(fill="Number of datasets")
g1
# ggsave(filename = "Fig6_supp.tiff", plot = g1, device = "tiff", path = "../paper_materials/plots_after_review/", dpi = 600, units = "in", width = 4, height = 5, scale = 2, compression = "lzw")
```



```{r fig.height=12}
g1 <- ggplot(df_hits_last_iter_agg, aes(x=cancer_type, y=feature,fill=n_best)) + 
  geom_tile() + 
  scale_fill_gradientn(colours = pal) + 
  xlab("Cancer type") + 
  ylab("Feature") +
  theme(
    text = element_text(size = 8, family = "arial"),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    ) + 
  labs(fill = "Number of \nexecutions")
g1
```



Aggregation by feature group for each cancer type: number of datasets in which features of the feature group selected.


```{r fig.width=10}
df_hits_last_iter_agg_new <- df_hits_last_iter %>%
  group_by(cancer_type, feature_group) %>%
  summarize(
    n_best = n_distinct(resample, repeat.)
    )

df_hits_last_iter_agg_new <- df_hits_last_iter_agg_new %>%
  inner_join(
    proper_feature_group_names,
    by=c("feature_group" = "feature_group_name")
  )

ggplot(df_hits_last_iter_agg_new, aes(x=cancer_type, y=n_best, fill=feature_group_name_proper)) +
  geom_bar(stat="identity", color="black", position="dodge") + 
  theme_light_custom()
```

Выводы: 

По типу рака:
В более половины датасетов встретились признаки следующих групп:

 - blood - вторичные структуры + регионы
 - bone - вторичные структуры + регионы + транскрипционные факторы
 - brain - вторичные структуры
 - breast - вторичные структуры + регионы + транскрипционные факторы + хроматин + метилирование
 - liver -  транскрипционные факторы
 - ovary - вторичные структуры + регионы + транскрипционные факторы
 - pancreatic -  регионы + транскрипционные факторы
 - prostate - вторичные структуры + регионы + транскрипционные факторы
 - skin - вторичные структуры + регионы + транскрипционные факторы
 - uterus - вторичные структуры + транскрипционные факторы + хроматин

Больше всего важных групп фичей наблюдается для рака груди, меньше всего - для мозга и печени.

По признакам:
В целом, нет группы фичей, которая важна для всех типов рака, однако для 8 типов рака важны вторичные структуры и транскрипционные факторы.

В рисунок по фичам не вошли признаки, которые ни разу не появились в финальном наборе фичей.
```{r}
all_f <- unique(df$feature)
all_f <- setdiff(all_f, grep(x = all_f, pattern = "shadow", value = T))

sel_f <- unique(df_hits_last_iter$feature)

setdiff(all_f, sel_f)
```
Это 9 фичей: модификации гистонов дли blood, brain, DnaseSeq для liver.

Алгоритм Борута был применен всего к 10 * 30 * 10 = 3000 датасетам. Посмотрим, какое количество раз каждый признак был включен в итоговый набор фичей в этих датасетах.
```{r}
by_f <- df_hits_last_iter %>%
  group_by(feature_group, feature) %>%
  summarize(
    total_sets = n()
    ) %>%
  arrange(desc(total_sets))
by_f
```

Абсолютный максимум у признака upper_Direct_Repeat - значение покрытия вторичной структурой direct repeat на уровне 1 Mb (908 раз). Далее, более 500 раз были использованы признаки  upper_G_quadruplex, G_quadruplex, upper_cancer_liver_SP1.human, Direct_Repeat, upper_Z_DNA_Motif, upper_cancer_liver_RXRA.human, upper_X5UTR, Short_Tandem_Repeat,	upper_cancer_liver_NR2F2.human,  upper_Short_Tandem_Repeat (то есть по большей части вторичные структуры). При этом в 8 из 11 случаев это значение покрытия на уровне 1 Мб.

```{r fig.width=10, fig.height=12}
ggplot(by_f, aes(x=reorder(feature, total_sets, median), y=total_sets, fill=feature_group))+
  geom_bar(stat="identity",color="black") + 
  coord_flip() + 
  scale_fill_jco()
```
```{r fig.width=10, fig.height=7}
by_f$is_upper <- 0
by_f$is_upper[grep(x=by_f$feature, pattern = "upper_")] <- 1
by_f$new_feature<- by_f$feature
by_f$new_feature[grep(x=by_f$new_feature, pattern = "DNase_seq")] <- "DNase_seq"
by_f$new_feature[(by_f$new_feature == "DNase_seq") & (by_f$is_upper == 1)] <- "upper_DNase_seq" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "DNA_methylation")] <- "DNA_methylation"
by_f$new_feature[(by_f$new_feature == "DNA_methylation") & (by_f$is_upper == 1)] <- "upper_DNA_methylation" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K4me3")] <- "H3K4me3"
by_f$new_feature[(by_f$new_feature == "H3K4me3") & (by_f$is_upper == 1)] <- "upper_H3K4me3"
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K36me3")] <- "H3K36me3"
by_f$new_feature[(by_f$new_feature == "H3K36me3") & (by_f$is_upper == 1)] <- "upper_H3K36me3" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K27ac")] <- "H3K27ac"
by_f$new_feature[(by_f$new_feature == "H3K27ac") & (by_f$is_upper == 1)] <- "upper_H3K27ac" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K27me3")] <- "H3K27me3"
by_f$new_feature[(by_f$new_feature == "H3K27me3") & (by_f$is_upper == 1)] <- "upper_H3K27me3" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K4me1")] <- "H3K4me1"
by_f$new_feature[(by_f$new_feature == "H3K4me1") & (by_f$is_upper == 1)] <- "upper_H3K4me1" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K9me3")] <- "H3K9me3"
by_f$new_feature[(by_f$new_feature == "H3K9me3") & (by_f$is_upper == 1)] <- "upper_H3K9me3" 
by_f$new_feature[grep(x=by_f$new_feature, pattern = "H3K9me3")] <- "H3K9me3"
by_f$new_feature[(by_f$new_feature == "H3K9me3") & (by_f$is_upper == 1)] <- "upper_H3K9me3"
by_f$new_feature[by_f$feature_group == "tf"] <- gsub(x = by_f$new_feature[by_f$feature_group == "tf"], pattern = "cancer_liver_", "")
by_f$new_feature[by_f$feature_group == "tf"] <- gsub(x = by_f$new_feature[by_f$feature_group == "tf"], pattern = ".human", "")

feat_colors <- as.character(proper_feature_group_names$colors)
names(feat_colors) <- as.character(proper_feature_group_names$feature_group_name_proper)

by_f <- by_f %>%
  # inner_join(df_feat_proper, by="feature") %>%
  inner_join(proper_feature_group_names, by=c("feature_group"="feature_group_name"))

by_f <- by_f %>%  
  group_by(feature_group_name_proper, new_feature) %>%
  summarize(
    total_sets = sum(total_sets)
    )

g1 <- ggplot(by_f, aes(x=reorder(new_feature, total_sets, median), y=total_sets, fill=feature_group_name_proper))+
  geom_bar(stat="identity",color="black", position="dodge") + 
  coord_flip() + 
  scale_fill_manual(values = feat_colors)+
  theme_light_custom(scale_fill = F)+
  ylab("Total number of datasets")+
  xlab("Feature")+
  labs(fill='Feature group')
g1
# ggsave(filename = "Fig19b.tiff", plot = g1, device = "tiff", path = "../paper_materials/", dpi = 600, units = "in", width = 5, height = 3, scale = 2.2, compression = "lzw")

g2 <- ggplot(by_f[by_f$total_sets >= 300, ], aes(x=reorder(new_feature, total_sets, median), y=total_sets, fill=feature_group_name_proper))+
  geom_bar(stat="identity",color="black", position="dodge") + 
  coord_flip() + 
  scale_fill_manual(values = feat_colors)+
  theme_light_custom(scale_fill = F)+
  ylab("Total number of datasets")+
  xlab("Feature")+
  labs(fill='Feature group')

# ggsave(filename = "Fig7_supp.tiff", plot = g2, device = "tiff", path = "../paper_materials/plots_after_review/", dpi = 600, units = "in", width = 5, height = 3, scale = 2.2, compression = "lzw")
g2
```


```{r}
by_f_canc <- df_hits_last_iter %>%
  group_by(cancer_type, feature_group, feature) %>%
  summarize(
    total_sets = n()
    ) %>%
  arrange(desc(total_sets))

by_f_canc
```

```{r}
by_f_canc$is_upper <- 0
by_f_canc$is_upper[grep(x=by_f_canc$feature, pattern = "upper_")] <- 1
by_f_canc$new_feature<- by_f_canc$feature
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "DNase_seq")] <- "DNase_seq"
by_f_canc$new_feature[(by_f_canc$new_feature == "DNase_seq") & (by_f_canc$is_upper == 1)] <- "upper_DNase_seq" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "DNA_methylation")] <- "DNA_methylation"
by_f_canc$new_feature[(by_f_canc$new_feature == "DNA_methylation") & (by_f_canc$is_upper == 1)] <- "upper_DNA_methylation" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K4me3")] <- "H3K4me3"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K4me3") & (by_f_canc$is_upper == 1)] <- "upper_H3K4me3"
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K36me3")] <- "H3K36me3"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K36me3") & (by_f_canc$is_upper == 1)] <- "upper_H3K36me3" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K27ac")] <- "H3K27ac"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K27ac") & (by_f_canc$is_upper == 1)] <- "upper_H3K27ac" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K27me3")] <- "H3K27me3"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K27me3") & (by_f_canc$is_upper == 1)] <- "upper_H3K27me3" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K4me1")] <- "H3K4me1"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K4me1") & (by_f_canc$is_upper == 1)] <- "upper_H3K4me1" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K9me3")] <- "H3K9me3"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K9me3") & (by_f_canc$is_upper == 1)] <- "upper_H3K9me3" 
by_f_canc$new_feature[grep(x=by_f_canc$new_feature, pattern = "H3K9me3")] <- "H3K9me3"
by_f_canc$new_feature[(by_f_canc$new_feature == "H3K9me3") & (by_f_canc$is_upper == 1)] <- "upper_H3K9me3" 
by_f_canc <- by_f_canc %>%
  select(cancer_type, feature_group, new_feature, total_sets) %>%
  dcast(feature_group + new_feature ~ cancer_type, value.var = "total_sets") %>%
  melt(id.var=c("new_feature", "feature_group")) %>%
  mutate(
    total_sets = ifelse(is.na(value), 0 , value)
  )

by_f_canc_agg <- by_f_canc %>%
  group_by(feature_group, new_feature) %>%
  summarize(
    mean_sets = mean(total_sets),
    med_sets = median(total_sets)
  )

by_f_canc_agg %>%
  arrange(desc(med_sets))

feat_colors <- as.character(proper_feature_group_names$colors)
names(feat_colors) <- as.character(proper_feature_group_names$feature_group_name_proper)

by_f_canc_agg <- by_f_canc_agg %>%
  inner_join(proper_feature_group_names, by=c("feature_group"="feature_group_name"))

by_f_canc_agg$new_feature[by_f_canc_agg$feature_group == "tf"] <- gsub(x = by_f_canc_agg$new_feature[by_f_canc_agg$feature_group == "tf"], pattern = "cancer_liver_", "")
by_f_canc_agg$new_feature[by_f_canc_agg$feature_group == "tf"] <- gsub(x = by_f_canc_agg$new_feature[by_f_canc_agg$feature_group == "tf"], pattern = ".human", "")

by_f_canc <- by_f_canc %>%
  inner_join(df_feat_proper %>% select(-feature_group), by=c("new_feature"="feature")) %>% 
  inner_join(proper_feature_group_names, by=c("feature_group"="feature_group_name"))

by_f_canc$new_feature[by_f_canc$feature_group == "tf"] <- gsub(x = by_f_canc$new_feature[by_f_canc$feature_group == "tf"], pattern = "cancer_liver_", "")
by_f_canc$new_feature[by_f_canc$feature_group == "tf"] <- gsub(x = by_f_canc$new_feature[by_f_canc$feature_group == "tf"], pattern = ".human", "")

```


```{r fig.height=7}
g3 <- ggplot(by_f_canc_agg[by_f_canc_agg$med_sets >= 10, ], aes(x=reorder(new_feature, med_sets), y=med_sets, fill=feature_group_name_proper))+
  geom_bar(stat="identity", color="black", position="dodge")+
  coord_flip()+
  scale_fill_manual(values = feat_colors)+
  theme_light_custom(scale_fill = F)+
  xlab("Median number of datasets")+
  ylab("Feature")+
  labs(fill='Feature group')
g3
# ggsave(filename = "Fig19_med.tiff", plot = g3, device = "tiff", path = "../paper_materials/", dpi = 600, units = "in", width = 5, height = 3, scale = 2.2, compression = "lzw")
```



```{r}
best_f <- by_f_canc_agg %>%
  ungroup() %>%
  # arrange(desc(med_sets)) %>%
  arrange(desc(mean_sets)) %>%
  filter(row_number() <= 8) %>%
  select(new_feature) %>%
  as.vector()
best_f

by_f_canc_plot <- by_f_canc %>% inner_join(best_f)
by_f_canc_plot$new_feature[by_f_canc_plot$new_feature == "upper_cancer_liver_RXRA.human"] <- "upper_RXRA"
by_f_canc_plot$new_feature[by_f_canc_plot$new_feature == "upper_cancer_liver_CTCF.human"] <- "upper_CTCF"

g1 <- ggplot(by_f_canc_plot, aes(x=variable, y=total_sets, color=new_feature, group=new_feature))+
  geom_line()+
  geom_point()+
  theme_light_custom()+
  scale_color_jco()+
  xlab("Cancer type")+
  ylab("Total number of datasets")+
  labs(color="Feature")
g1
# ggsave(filename = "Fig8_supp.tiff", plot = g1, device = "tiff", path = "../paper_materials/plots_after_review/", dpi = 600, units = "in", width = 5, height = 3, scale = 2.2, compression = "lzw")
```

```{r}
# g2 <- ggplot(
#   by_f_canc_plot, aes(y=total_sets, x=factor(total_sets), fill=new_feature)
# )+
#   facet_wrap(~variable, scales="free", nrow=2)+
#   geom_bar(stat="identity", position="dodge", color="black")+
#   theme_light_custom()+
#   xlab("Feature")+
#   ylab("Total number of datasets")+
#   labs(fill="Feature")+
#   theme(
#         axis.text.x = element_blank(),
#         axis.ticks.x = element_blank()
#     )
# g2
# ggsave(filename = "Fig16.tiff", plot = g2, device = "tiff", path = "../paper_materials/", dpi = 600, units = "in", width = 5, height = 3, scale = 2.2, compression = "lzw")
```

By sum
```{r}
best_f <- by_f %>%
  arrange(desc(total_sets)) %>%
  ungroup() %>%
  filter(row_number() <= 8) %>%
  select(new_feature) %>%
  as.vector()
best_f
by_f_canc_plot <- by_f_canc %>% inner_join(best_f)
# by_f_canc_plot$new_feature[by_f_canc_plot$new_feature == "upper_cancer_liver_RXRA.human"] <- "upper_RXRA" 
# by_f_canc_plot$new_feature[by_f_canc_plot$new_feature == "upper_cancer_liver_CTCF.human"] <- "upper_CTCF" 

g1 <- ggplot(by_f_canc_plot, aes(x=variable, y=total_sets, color=new_feature, group=new_feature))+
  geom_line()+
  geom_point()+
  theme_light_custom()+
  scale_color_jco()+
  xlab("Cancer type")+
  ylab("Total number of datasets")+
  labs(color="Feature")
g1
# ggsave(filename = "Fig8_supp.tiff", plot = g1, device = "tiff", path = "../paper_materials/plots_after_review/", dpi = 600, units = "in", width = 5, height = 3, scale = 2.2, compression = "lzw")
```




```{r}
by_f_canc_out <- by_f_canc %>%
  inner_join(proper_feature_group_names, by=c("feature_group"="feature_group_name"))
by_f_canc_out$feature_name <- by_f_canc_out$feature
by_f_canc_out$feature_name[(by_f_canc_out$feature_group == "tf")] <- gsub(x=by_f_canc_out$feature_name[(by_f_canc_out$feature_group == "tf")], pattern = "cancer_liver_", replacement = "")
by_f_canc_out$feature_name[(by_f_canc_out$feature_group == "tf")] <- gsub(x=by_f_canc_out$feature_name[(by_f_canc_out$feature_group == "tf")], pattern = ".human", replacement = "")
# unique(by_f_canc_out$feature_name)[order(unique(by_f_canc_out$feature_name))]
by_f_canc_out <- by_f_canc_out %>%
  ungroup() %>%
  select(cancer_type, feature_group_name_proper, feature_name, total_sets) %>%
  arrange(desc(total_sets))
# write.csv(by_f_canc_out, "../paper_materials/table_boruta.csv")
```



Теперь выберем пул фичей для каждого типа рака.

```{r,fig.width=12, fig.height=5}
ggplot(df_hits_last_iter_agg, aes(n_best)) + 
  geom_density(alpha=0.5) + 
  facet_grid(~cancer_type, scales = "free")
```

Видно, что 50 - для многих типов рака переходная граница между большим количеством фичей около 0 и небольшим количеством важных фичей. Выберем эти число как порог для выбора фичей.  

```{r}
thr <- 50
df_hits_last_iter_agg %>%
  filter(n_best > 50) %>%
  group_by(cancer_type) %>%
  summarize(
    n_features = n()
  ) %>%
  arrange(n_features)
```

```{r}
res_hsp <- df_hits_last_iter_agg %>%
  mutate(boruta_selected = ifelse(n_best > thr, 1, 0)) %>%
  select(-thr)

res_hsp %>%  
  filter(boruta_selected == 1) %>%
  ungroup() %>%
  select(feature) %>%
  unique() %>%
  nrow()
```

Всего 46 из 130 фичей

```{r fig.width=10}
cn_for_feat <- res_hsp %>% 
  filter(boruta_selected == 1) %>%
  group_by(feature, feature_group) %>%
  summarize(n_cancer_types = n_distinct(cancer_type)) %>% 
  arrange(desc(n_cancer_types))

ggplot(cn_for_feat, aes(x=reorder(feature, n_cancer_types, mean) , y=n_cancer_types,fill=feature_group))+
  geom_bar(stat="identity",color="black",  position="dodge") + 
  coord_flip() + 
  scale_fill_jco()
```

Heatmap for selected features

```{r fig.width=10}
ggplot(res_hsp %>% filter(boruta_selected == 1), aes(x=cancer_type, y=feature,fill=n_best)) + 
  geom_tile() + 
  scale_fill_gradientn(colours = pal)
```



```{r save}
# wb <- createWorkbook()
# addWorksheet(wb, "hotspots_features")
# writeData(wb, sheet="hotspots_features", res_hsp)
# saveWorkbook(wb, "../data/output/boruta_stats.xlsx", overwrite = T)
```


### Cancer types clustering


```{r fig.width=15}
df_hits_last_iter_agg$feature_n <- gsub(df_hits_last_iter_agg$feature, pattern = "upper_", replacement = "")
# specific preparation for tissue-specific features
df_hits_last_iter_agg$feature_n[df_hits_last_iter_agg$feature_group == "histones"] <- unlist(
  lapply(
    strsplit(
      x = df_hits_last_iter_agg$feature_n[df_hits_last_iter_agg$feature_group == "histones"],
      split="_"), 
    function(x) x[[3]])
  )
df_hits_last_iter_agg$feature_n[df_hits_last_iter_agg$feature_group == "chromatin"] <- "DNase_seq"


# perform clustering for each feature group separately
f_gr <- unique(df_hits_last_iter_agg$feature_group)
f_gr <- f_gr[f_gr != "tad"]

all_cl_res <- data.frame()

for (feat_gr_nm  in f_gr){
  data_part <- df_hits_last_iter_agg[df_hits_last_iter_agg$feature_group == feat_gr_nm,
                                     c("cancer_type", "feature_n", "n_best")]
  data_part_max <- data_part %>%
    group_by(cancer_type, feature_n) %>%
    filter(n_best == max(n_best)) %>%
    mutate(
      rn = row_number()
    ) %>%
    filter(rn == 1)
  
  data_part_max_casted <- dcast(data_part_max, cancer_type ~ feature_n, value.var = "n_best")
  data_part_max_casted[is.na(data_part_max_casted)] <- 0
  row.names(data_part_max_casted) <- data_part_max_casted$cancer_type
  data_part_max_casted$cancer_type <- NULL
  dist_m <- dist(data_part_max_casted, method = "euclidean")
  hclustering <- hclust(dist_m, method = "ward.D")
  # ordering <- row.names(data_part_max_casted)[hclustering$order]
  # cl <- data.frame(
  #   feature_group = feat_gr_nm,
  #   cancer_type = ordering,
  #   cluster_n = seq(1, length(ordering))
  # )
  cl <- data.frame(
    feature_group = feat_gr_nm,
    cancer_type = row.names(data_part_max_casted),
    cluster_n = cutree(hclustering, k = 3)
  )
  all_cl_res <- rbind(all_cl_res, cl)
}

ggplot(all_cl_res, aes(x=feature_group, y=cluster_n, color=cancer_type, group=cancer_type)) + 
  geom_point() + 
  geom_line() +
  coord_polar() +
  theme(legend.position = "bottom") +
  geom_ribbon(aes(ymin=0, ymax=cluster_n, fill=cancer_type, alpha=0.2)) + 
  facet_wrap( ~ cancer_type, nrow=2, ncol=5) + 
  scale_fill_jco() + 
  scale_color_jco()

```

Нет идентичных типов рака, однако некоторые типы рака выглядят похоже, например:

 - blood / pancreatic
 - bone / brain
 - ovary / prostate (!)
 - breast / uterus (!)

```{r}
all_cl_raw <- dcast(all_cl_res, cancer_type ~ feature_group, value.var = "cluster_n")
all_cl_raw
```



```{r echo=F}
# FInal clustering - NOT STABLE
# in_m <- as.matrix(all_cl_raw)
# n <- nrow(all_cl_raw)
# out_m <- matrix(nrow=n, ncol=n)
# out_m
# for (i in 1: (n-1)){
#   for (j in (i + 1): n){
#     x1  <- in_m[i, ]
#     x2  <- in_m[j, ]
#     d <- (x1 - x2)^2
#     mean_d <- sum(d[!is.na(d)]) / length(d[!is.na(d)])
#     out_m[i, j] <- out_m[j, i] <- mean_d     
#   }
#  
# }
# rownames(out_m) <- row.names(all_cl_raw)
# colnames(out_m) <- row.names(all_cl_raw)
# dist_m <- as.dist(out_m)
# dist_m
# # dist_m <- dist(all_cl_raw, method = "manhattan")
# hclustering <- hclust(dist_m, method = "ward.D2")
# ordering <- row.names(all_cl_raw)[hclustering$order]
# final_cl <- data.frame(
#   cancer_type = ordering,
#   cluster_n = seq(1, length(ordering))
# )
# plot(hclustering$height)
# plot(hclustering)
# rect.hclust(hclustering, k=4)
```


### Feature correlation

```{r fig.width=10}
best_features <- read.xlsx("../data/output/boruta_stats.xlsx", sheet = "hotspots_features")
best_features <- best_features[best_features$boruta_selected == 1, 
                               c("cancer_type", "feature")]
best_features <- best_features %>% select(feature) %>% unique()
features_cols <- gsub(x=best_features$feature, pattern = "upper_", replacement = "") %>% unique()

# load data
win_len <- 100000
win_len_upper <- win_len * 10
data_path <- "../data/datasets/" 
data <- read.csv(
  paste0(data_path, "dataset_", format(win_len, scientific = FALSE), ".csv")
)
source("features.R")
all_data <- get_higher_level_features(data=data, features_cols = features_cols, 
                                           win_len_upper = win_len_upper,
                                           path_to_upper_data = paste0(data_path, "dataset_",
                                                                       format(win_len_upper, scientific = FALSE), ".csv"))

corrm <- cor(all_data[best_features$feature], method="spearman")
corrm_m <- melt(corrm)

corrm_m %>%
  filter(Var1 != Var2) %>%
  arrange(desc(value)) %>%
  head(15)
corrm_m$Var1 <- factor(corrm_m$Var1, levels = best_features$feature %>% sort())
corrm_m$Var2 <- factor(corrm_m$Var2, levels = best_features$feature %>% sort())
ggplot(corrm_m , aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + 
  scale_fill_gradientn(colours = pal) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
ggplot(corrm_m %>% filter(Var1 != Var2), aes(value)) + 
  geom_histogram()
```

```{r}
corrm_m %>% 
  filter(Var1 != Var2) %>%
  filter(value >= 0.8) %>%
  arrange(desc(value))

```

