---
title: "Model selection analysis"
output: html_notebook
---

```{r echo=F}
library(ggplot2)
library(ggsci)
library(wesanderson)

library(dplyr)
```


### Random Forest without winsorizing

```{r}
all_results <- read.csv("../data/output/model_selection/hp_rf.csv")
all_results$diff_auc <- all_results$tr_auc - all_results$te_auc
part_res <- read.csv("../data/output/model_selection/hp_rf_0.csv")
all_results <- rbind(all_results, part_res)


res_all <- all_results %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    q30 = quantile(te_auc, 0.3),
    sd = sd(te_auc)
  ) %>%
  arrange(cancer_type)

res_all %>%
  filter(min_te_auc > 0.55)

ggplot(all_results, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)

```

Select optimal parameters on good cancer types

```{r}
good_cancers_nm <- c("breast", "prostate", "ovary")
pred_cancers_data <- all_results %>%
  filter(cancer_type %in% good_cancers_nm)

# all_results %>%
#   filter(cancer_type %in% good_cancers_nm) %>%
#   group_by(ntree) %>%
#   summarize(
#     min_te_auc = min(te_auc),
#     q10 = quantile(te_auc, 0.1),
#     q30 = quantile(te_auc, 0.3),
#     q70 = quantile(te_auc, 0.7),
#     max_te = max(te_auc)
#   ) %>%
#   arrange(min_te_auc)

gb_pars <- pred_cancers_data %>%
  group_by(ntree, nodesize, maxnodes, mtry, cancer_type) %>%
  summarize(
    mean_d = mean(diff_auc),
    med = median(diff_auc),
    sd = sd(te_auc)
  ) %>%
  group_by(cancer_type) %>%
  mutate(
    mean_rank = dense_rank(mean_d),
    med_rank = dense_rank(med),
    sd_rank = ntile(sd, 5)
  )

gb_pars1 <- gb_pars %>%
  group_by(ntree, nodesize, maxnodes, mtry) %>%
  summarize(
    mean_rank = mean(mean_rank),
    med_rank = mean(med_rank),
    sd_rank = mean(sd_rank)
  ) %>%
  mutate(summed_rank = mean_rank + med_rank + sd_rank)

par_opt <- gb_pars1 %>%
  ungroup() %>%
  filter(summed_rank == min(summed_rank))

par_opt
```

Stats of models with selected parameters

```{r}
opt_pars <- all_results %>%
  inner_join(par_opt, by=c("nodesize", "ntree", "maxnodes", "mtry"))

res_opt <- opt_pars %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    med_te_auc = median(te_auc),
    mean_te_auc = mean(te_auc),
    q80 = quantile(te_auc, 0.8),
    q90 = quantile(te_auc, 0.9),
    max_te_auc = max(te_auc),
    sd = sd(te_auc)
  ) %>%
  arrange(cancer_type)

ggplot(opt_pars, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)

```


### Random Forest with winsorizing

```{r}
all_results <- read.csv("../data/output/model_selection/hp_rf_lo.csv")
all_results$diff_auc <- all_results$tr_auc - all_results$te_auc
part_res <- read.csv("../data/output/model_selection/hp_rf_lo_0.csv")
all_results <- rbind(all_results, part_res)


res_all <- all_results %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    q30 = quantile(te_auc, 0.3),
    sd = sd(te_auc)
  ) 
res_all$cancer_type <- as.character(res_all$cancer_type)
res_all <- res_all %>%
  arrange(cancer_type)


res_all %>%
  filter(min_te_auc > 0.55)

ggplot(all_results, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)

```

Select optimal parameters on good cancer types

```{r}
good_cancers_nm <- c("breast", "prostate", "ovary")
pred_cancers_data <- all_results %>%
  filter(cancer_type %in% good_cancers_nm)

# all_results %>%
#   filter(cancer_type %in% good_cancers_nm) %>%
#   group_by(ntree) %>%
#   summarize(
#     min_te_auc = min(te_auc),
#     q10 = quantile(te_auc, 0.1),
#     q30 = quantile(te_auc, 0.3),
#     q70 = quantile(te_auc, 0.7),
#     max_te = max(te_auc)
#   ) %>%
#   arrange(min_te_auc)

gb_pars <- pred_cancers_data %>%
  group_by(ntree, nodesize, maxnodes, mtry, cancer_type) %>%
  summarize(
    mean_d = mean(diff_auc),
    med = median(diff_auc),
    sd = sd(te_auc)
  ) %>%
  group_by(cancer_type) %>%
  mutate(
    mean_rank = dense_rank(mean_d),
    med_rank = dense_rank(med),
    sd_rank = ntile(sd, 5)
  )

gb_pars <- gb_pars %>%
  group_by(ntree, nodesize, maxnodes, mtry) %>%
  summarize(
    mean_rank = mean(mean_rank),
    med_rank = mean(med_rank),
    sd_rank = mean(sd_rank)
  ) %>%
  mutate(summed_rank = mean_rank + med_rank + sd_rank)

par_opt <- gb_pars %>%
  ungroup() %>%
  filter(summed_rank == min(summed_rank))

par_opt
```

Stats of models with selected parameters

```{r}
opt_pars <- all_results %>%
  inner_join(par_opt, by=c("nodesize", "ntree", "maxnodes", "mtry"))

res_opt <- opt_pars %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    med_te_auc = median(te_auc),
    mean_te_auc = mean(te_auc),
    q80 = quantile(te_auc, 0.8),
    q90 = quantile(te_auc, 0.9),
    max_te_auc = max(te_auc),
    sd = sd(te_auc)
  )
res_opt$cancer_type <- as.character(res_opt$cancer_type)
res_opt <- res_opt %>%
  arrange(cancer_type)

ggplot(opt_pars, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)

```

### PLS + Random Forest without winsorizing


```{r}
all_results <- read.csv("../data/output/model_selection/hp_rf_pls_lo.csv")
all_results$diff_auc <- all_results$tr_auc - all_results$te_auc
res_all <- all_results %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    q30 = quantile(te_auc, 0.3),
    sd = sd(te_auc)
  ) %>%
  arrange(cancer_type)

# make decision on alll cancer types
pred_cancers_data <- all_results

ggplot(pred_cancers_data, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)
```

Select optimal parameters

```{r}
gb_pars <- pred_cancers_data %>%
  group_by(ntree, nodesize, maxnodes, mtry, ncomp, cancer_type) %>%
  summarize(
    mean_d = mean(diff_auc),
    med = median(diff_auc),
    sd = sd(te_auc)
  ) %>%
  group_by(cancer_type) %>%
  mutate(
    mean_rank = dense_rank(mean_d),
    med_rank = dense_rank(med),
    sd_rank = ntile(sd, 5)
  )
gb_pars <- gb_pars %>%
  group_by(ntree, nodesize, maxnodes, mtry, ncomp) %>%
  summarize(
    mean_rank = mean(mean_rank),
    med_rank = mean(med_rank),
    sd_rank = mean(sd_rank)
  ) %>%
  mutate(summed_rank = mean_rank + med_rank + sd_rank)

par_opt <- gb_pars %>%
  ungroup() %>%
  filter(summed_rank == min(summed_rank))
par_opt
```
Stats of models with selected parameters

```{r}

opt_pars <- all_results %>%
  inner_join(par_opt, by=c("nodesize", "ntree", "maxnodes", "mtry", "ncomp"))

res_opt <- opt_pars %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    med_te_auc = median(te_auc),
    mean_te_auc = mean(te_auc),
    q80 = quantile(te_auc, 0.8),
    q90 = quantile(te_auc, 0.9),
    max_te_auc = max(te_auc),
    sd = sd(te_auc)
  ) %>%
  arrange(cancer_type)

ggplot(opt_pars, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)
```


### PCA + Random Forest without winsorizing


```{r}
all_results <- read.csv("../data/output/model_selection/hp_rf_pca_lo.csv")
all_results$diff_auc <- all_results$tr_auc - all_results$te_auc
res_all <- all_results %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    q30 = quantile(te_auc, 0.3),
    sd = sd(te_auc)
  ) %>%
  arrange(cancer_type)

ggplot(all_results, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)
```

Select optimal parameters

```{r}
good_cancers_nm <- c("prostate", "ovary", "breast")
pred_cancers_data <- all_results %>%
  filter(cancer_type %in% good_cancers_nm)

gb_pars <- pred_cancers_data %>%
  group_by(ntree, nodesize, maxnodes, mtry, pcaComp, cancer_type) %>%
  summarize(
    mean_d = mean(diff_auc),
    med = median(diff_auc),
    sd = sd(te_auc)
  ) %>%
  group_by(cancer_type) %>%
  mutate(
    mean_rank = dense_rank(mean_d),
    med_rank = dense_rank(med),
    sd_rank = ntile(sd, 5)
  )
gb_pars <- gb_pars %>%
  group_by(ntree, nodesize, maxnodes, mtry, pcaComp) %>%
  summarize(
    mean_rank = mean(mean_rank),
    med_rank = mean(med_rank),
    sd_rank = mean(sd_rank)
  ) %>%
  mutate(summed_rank = mean_rank + med_rank + sd_rank)

par_opt <- gb_pars %>%
  ungroup() %>%
  filter(summed_rank == min(summed_rank))
par_opt
```

```{r}
opt_pars <- all_results %>%
  inner_join(par_opt, by=c("ntree", "nodesize", "maxnodes", "mtry", "pcaComp"))

res_opt <- opt_pars %>%
  group_by(cancer_type) %>%
  summarize(
    min_te_auc = min(te_auc),
    q10 = quantile(te_auc, 0.1),
    q20 = quantile(te_auc, 0.2),
    med_te_auc = median(te_auc),
    mean_te_auc = mean(te_auc),
    q80 = quantile(te_auc, 0.8),
    q90 = quantile(te_auc, 0.9),
    max_te_auc = max(te_auc),
    sd = sd(te_auc)
  ) %>%
  arrange(cancer_type)

ggplot(opt_pars, aes(te_auc)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type)
```


### Compare PCA RF and RF
```{r}
opt_pars_pca <- opt_pars %>%
  mutate(model = 'pca rf')

all_results <- read.csv("../data/output/model_selection/hp_rf.csv")
all_results$diff_auc <- all_results$tr_auc - all_results$te_auc
part_res <- read.csv("../data/output/model_selection/hp_rf_0.csv")
all_results <- rbind(all_results, part_res)

opt_pars_rf <- all_results %>%
  filter(
    ntree == 500,
    mtry == 5,
    maxnodes == 3,
    nodesize == 30
    ) %>%
  mutate(model = 'rf')

opt_pars_comp <- rbind(
  opt_pars_pca %>% select(cancer_type, te_auc, model),
  opt_pars_rf %>% select(cancer_type, te_auc, model)
  )

ggplot(opt_pars_comp, aes(te_auc, fill = model)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ cancer_type) + scale_fill_jco()
```


```{r}
opt_pars_comp %>%
  group_by(cancer_type,model) %>%
  summarize(
    min(te_auc),
    median(te_auc)
  )
```

