---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(reshape2)

library(ggplot2)
library(wesanderson)
library(ggsci)
library(ggpubr)

library(openxlsx)

library(formattable)

customYellow = 'yellow'
customBlue = '#1E90FF'
pal <- wes_palette("Zissou1", 100, type = "continuous")

source('../run/tools.R')
source('../run/features.R')

colors <- as.character(proper_feature_group_names$colors)
names(colors) <- as.character(proper_feature_group_names$feature_group_name_proper)

# load new data
data_path <- "../data/output/stepwise_addition/"
fs <- read.csv(paste0(data_path, "result_features_selected_100000.csv"), row.names = 1, stringsAsFactors = F)
fs0 <- read.csv(paste0(data_path, "result_features_selected_100000_0.csv"), row.names = 1, stringsAsFactors = F)
fs <- rbind(fs0, fs)
fs$agg_level[fs$agg_level == "99"] <- "99."

roc_auc <- read.csv(paste0(data_path, "result_roc_auc_100000.csv"), row.names = 1, stringsAsFactors = F)
roc_auc0 <- read.csv(paste0(data_path, "result_roc_auc_100000_0.csv"), row.names = 1, stringsAsFactors = F)
roc_auc <- rbind(roc_auc0, roc_auc)
roc_auc$agg_level[roc_auc$agg_level == "99"] <- "99."

recall <- read.csv(paste0(data_path, "result_recall_100000.csv"), row.names = 1, stringsAsFactors = F)
recall0 <- read.csv(paste0(data_path, "result_recall_100000_0.csv"), row.names = 1, stringsAsFactors = F)
recall <- rbind(recall0, recall)
recall$agg_level[recall$agg_level == "99"] <- "99."

# load baseline data


cor_canc <- c("brain",  "liver", "ovary", "prostate", "pancreatic")
roc_auc <- roc_auc[roc_auc$cancer_type %in% cor_canc, ]
recall <- recall[recall$cancer_type %in% cor_canc, ]
fs <- fs[fs$cancer_type %in% cor_canc, ]
```

Determine number of successfull iterations

```{r}
# join baseline: feature_selection_iter=0
fs <- fs %>%
  arrange(cancer_type, agg_level, win_len, feature_selection_iter) %>%
  group_by(cancer_type, agg_level, win_len) %>%
  mutate(
    lag_mean_lift_recall_003 = lag(mean_lift_recall_003),
    diff_mean_lift_recall_003 = mean_lift_recall_003 - lag_mean_lift_recall_003,
    ratio_up_mean_lift_recall_003 = diff_mean_lift_recall_003 / lag_mean_lift_recall_003,
    lag_mean_roc = lag(mean_te_auc),
    diff_mean_roc = mean_te_auc - lag_mean_roc
    ) %>%
  mutate(
    diff_up_greater_3 = ifelse(diff_mean_lift_recall_003 >= 0.1, 1, 0)
  )

fs_last_iter <- fs %>%
  filter(diff_up_greater_3 == 0) %>%
  group_by(cancer_type, agg_level, win_len) %>%
  filter(row_number() == 1) %>%
  mutate(last_feature_selection_iter = feature_selection_iter) %>%
  select(cancer_type, agg_level, win_len, last_feature_selection_iter)

fs <- fs %>%
  left_join(
    fs_last_iter, by=c("cancer_type", 'agg_level', "win_len")
  ) %>%
  mutate(
    last_feature_selection_iter = ifelse(is.na(last_feature_selection_iter), 5, last_feature_selection_iter)
  ) %>%
  filter(feature_selection_iter <= last_feature_selection_iter)
  
```

Select best iterations

```{r}
roc_auc_best <- fs %>%
  select("cancer_type","agg_level", "win_len", "feat_added", "feature_selection_iter") %>%
  inner_join(
    roc_auc, by=c("cancer_type","agg_level", "win_len", "feat_added", "feature_selection_iter"))

recall_best <- fs %>%
  select("cancer_type","agg_level", "win_len", "feat_added", "feature_selection_iter") %>%
  inner_join(
    recall, by=c("cancer_type","agg_level", "win_len", "feat_added", "feature_selection_iter"))
```

Summary stats for checking

```{r}
roc_auc_best_summary <- roc_auc_best %>%
  group_by(cancer_type,agg_level, win_len, feat_added, feature_selection_iter) %>%
  summarize(
    mean_test_auc = mean(te_auc),
    mean_train_auc = mean(tr_auc)
  )

map_df <- data.frame(agg_level=c("99.", "99.5.", "99.9."), agg_level_num = c(0.01, 0.005, 0.001))
recall_best <- recall_best %>%
  inner_join(map_df)
recall_best$lift_precision <- recall_best$precision / recall_best$agg_level_num
recall_best_summary <- recall_best %>%
  mutate() %>%
  group_by(cancer_type,agg_level, win_len, feat_added, feature_selection_iter, quantile) %>%
  summarize(
    mean_lift_recall = mean(lift_recall),
    mean_lift_precision = mean(lift_precision)
  )  
```

Distribution of number of successful iterations
```{r}
max_iter <- fs %>%
  group_by(cancer_type, agg_level, win_len) %>%
  filter(feature_selection_iter == max(feature_selection_iter))


ggplot(max_iter,
       aes(x=reorder(cancer_type, feature_selection_iter, FUN=mean), 
           y=feature_selection_iter,fill=agg_level)) +
  geom_bar(stat="identity",color="black", position="dodge")+ 
  theme_light()

```



```{r}
df_onegroup <- read.xlsx('../reports/summary14.xlsx')

df_onegroup <- df_onegroup %>%
  inner_join(
    proper_feature_group_names,
    by=c("Feature.group" = "feature_group_name")
  )
df_onegroup_copy <- df_onegroup
```



```{r}
best_feature_groups <- data.frame(

  cancer_type=c("blood", "bone", "brain", 'breast', "liver", "ovary", "prostate", "pancreatic", "skin", "uterus"),
  feature_group=c("histones", "sec_str", "sec_str", "tf", "tf", "tf", "tf", "reg", "tf", "chromatin"),
  feature_selection_iter=0
)
best_feature_groups <- best_feature_groups %>%
  left_join(
    df_onegroup %>% 
      select(Cancer.type, Labeling.type, Feature.group, Mean.lift.of.recall,
             Mean.test.ROC.AUC), 
    by=c('cancer_type'='Cancer.type', 'feature_group'='Feature.group')
  )
best_feature_groups <- best_feature_groups %>%
  mutate(
    win_len=100000
  ) %>%
  select(cancer_type, Labeling.type, win_len, feature_group, Mean.lift.of.recall,
             Mean.test.ROC.AUC, feature_selection_iter) %>%
  setNames(c("cancer_type","agg_level", "win_len", "feat_added", "mean_lift_recall_003", "mean_te_auc",
             "feature_selection_iter"))
best_feature_groups
```

```{r}
all_fs <- rbind.data.frame(
  fs %>%
  select(names(best_feature_groups)),
  best_feature_groups
)
all_fs <- all_fs %>%
  arrange(cancer_type, agg_level, win_len, feature_selection_iter) %>%
  group_by(cancer_type, agg_level, win_len) %>%
  mutate(
    lag_mean_lift_recall_003 = lag(mean_lift_recall_003),
    diff_mean_lift_recall_003 = mean_lift_recall_003 - lag_mean_lift_recall_003,
    lag_mean_roc = lag(mean_te_auc),
    diff_mean_roc = mean_te_auc - lag_mean_roc
    )

```

Uplift for successfull iterations in terms of lift of recall and ROC AUC by cancer type
```{r}
ggplot(all_fs %>% filter(feature_selection_iter > 0),
       aes(x=feature_selection_iter, y=diff_mean_lift_recall_003, colour=cancer_type)) +
  geom_line()+ 
  geom_point()+ 
  facet_wrap(~agg_level, nrow=3, scales="free")+
  theme_light()+
  scale_color_jco()
```


```{r}
ggplot(all_fs %>% filter(feature_selection_iter > 0),
       aes(x=feature_selection_iter, y=diff_mean_roc, colour=cancer_type)) +
  geom_line()+ 
  geom_point()+ 
  facet_wrap(~agg_level, nrow=3)+
  theme_light()+
  scale_color_jco()
```

```{r}
roc_auc_ <- max_iter %>%
  select(cancer_type, agg_level, win_len, feature_selection_iter, feat_added) %>%
  inner_join(
    roc_auc_best
  ) %>%
  mutate(
    train_test_diff = tr_auc - te_auc
  ) %>%
  group_by(cancer_type, agg_level) %>%
  summarize(
    min_test_auc = min(te_auc),
    sd_test_auc = sd(te_auc),
    median_test_auc = median(te_auc),
    mean_test_auc = mean(te_auc),
    train_test_diff = mean(train_test_diff),
    n_ex = n()
  ) %>%
  mutate(
    ci_lower_roc_auc = mean_test_auc - 1.96 * sd_test_auc / sqrt(n_ex),
    ci_upper_roc_auc = mean_test_auc + 1.96 * sd_test_auc / sqrt(n_ex), 
  ) %>%
  select(-n_ex)

recall_stats <- max_iter %>%
  select(cancer_type, agg_level, win_len, feature_selection_iter, feat_added) %>%
  inner_join(
    recall_best
  ) %>%
  group_by(cancer_type, agg_level, quantile) %>%
  summarize(
    mean_recall = mean(recall),
    mean_lift_recall = mean(lift_recall),
    mean_precision = mean(precision),
    mean_lift_precision = mean(lift_precision),
    sd_recall = sd(recall),
    sd_lift_recall = sd(lift_recall),
    sd_precision = sd(precision),
    sd_lift_precision = sd(lift_precision),
    n_ex = n()
  ) %>%
  mutate(
    ci_lower_recall = mean_recall - 1.96 * sd_recall / sqrt(n_ex),
    ci_upper_recall = mean_recall + 1.96 * sd_recall / sqrt(n_ex),
    
    ci_lower_lift_recall = mean_lift_recall - 1.96 * sd_lift_recall / sqrt(n_ex),
    ci_upper_lift_recall = mean_lift_recall + 1.96 * sd_lift_recall / sqrt(n_ex),    
    
    ci_lower_precision = mean_precision - 1.96 * sd_precision / sqrt(n_ex),
    ci_upper_precision = mean_precision + 1.96 * sd_precision / sqrt(n_ex),       
    
    ci_lower_lift_precision = mean_lift_precision - 1.96 * sd_lift_precision / sqrt(n_ex),
    ci_upper_lift_precision = mean_lift_precision + 1.96 * sd_lift_precision / sqrt(n_ex)
  )
```

Final comparison - BFS, One feature group and One feature group + Stepwise addition
Final conclusion: which types of cancer gained additional quality (rank by uplift cancer type/agg level)
Пересечение доверительных интервалов для новой модели и модели на 1 группе признаков


```{r}
onegr <- df_onegroup %>%
  inner_join(
    best_feature_groups %>% select(cancer_type, feat_added) %>% unique() %>% as.data.frame(),
    by=c("Cancer.type"="cancer_type", "Feature.group"="feat_added")
  ) %>%
  select("Cancer.type", "Labeling.type", "Lower.bound.of.confidence.interval.for.mean.lift.of.recall",
         "Mean.lift.of.recall", "Upper.bound.of.confidence.interval.for.mean.lift.of.recall",
         "Lower.bound.of.confidence.interval.for.mean.test.ROC.AUC","Mean.test.ROC.AUC",
         "Upper.bound.of.confidence.interval.for.mean.test.ROC.AUC"
         ) %>%
  setNames(c("cancer_type", "agg_level","ci_lower_lift_recall", "mean_lift_recall", "ci_upper_lift_recall",
             "ci_lower_roc_auc","mean_test_auc", "ci_upper_roc_auc"
             )) %>%
  mutate(
    model_type = "best_group"
  )
```



Mean lift of recall (0.03)
```{r fig.height=7}
recall_stats$model_type <- "stepwise_addition"
recall_stats_a <- recall_stats %>% filter(quantile == 0.03) %>% 
  select(c("cancer_type", "agg_level", "model_type","ci_lower_lift_recall", "mean_lift_recall", "ci_upper_lift_recall"))
all_d <- rbind.data.frame(
  recall_stats_a, 
  onegr %>% select(c("cancer_type", "agg_level", "model_type","ci_lower_lift_recall", "mean_lift_recall", "ci_upper_lift_recall")))

ggplot(all_d , aes(
  x=agg_level, y=mean_lift_recall, fill=model_type, ymin=ci_lower_lift_recall,
  ymax=ci_upper_lift_recall)
  ) + 
  geom_bar(position="dodge", color="black", stat="identity")+
  # geom_errorbar()+
  facet_wrap(~cancer_type, ncol=2, scales="free")+
  theme_light()
```

Mean test ROC AUC

```{r}
roc_auc_$model_type <- "stepwise_addition"
ggplot(roc_auc_ , aes(
  x=agg_level, y=mean_test_auc, fill=model_type, ymin=ci_lower_roc_auc,
  ymax=ci_upper_roc_auc)
  ) + 
  geom_bar(position="dodge", color="black", stat="identity")+
  geom_errorbar(width=0.2)+
  facet_wrap(~cancer_type, ncol=2, scales="free")+
  theme_light()

```

Какие признаки добавились для каждого типа рака (из каких групп)

```{r fig.height=10, fig.width=9}
feature_cols <- unique(roc_auc$feat_added)
df_f <- get_feature_df(feature_cols)
df_f <- df_f %>% 
  inner_join(
    proper_feature_group_names,
    by=c("feature_group" = "feature_group_name")
  )

feat_added_stats <- fs %>%
  inner_join(
    df_f, 
    by=c("feat_added" = "feature")
    ) %>%
  group_by(cancer_type, agg_level, feature_group_name_proper) %>%
  summarize(
    n_features_added = n()
  ) %>%
  mutate(
    group = 1
  ) %>%
  group_by(cancer_type, agg_level) %>%
  mutate(rank_feature_group = dense_rank(feature_group_name_proper))

groups <- data.frame(rank_feature_group = seq(1, 7), fake_col = 1)
old_f <- best_feature_groups %>%
  inner_join(df_f, by=c("feat_added"="feature_group")) %>%
  select(cancer_type, agg_level, feature_group_name_proper) %>%
  mutate(
    n_features_added = NA,
    group = 0,
    fake_col = 1
  ) %>%
  unique() %>%
  full_join(groups) %>%
  select(-fake_col)

feat_add <- rbind.data.frame(
  feat_added_stats,
  old_f
) 
feat_add$final_feat_group <- paste0(feat_add$agg_level, feat_add$rank_feature_group, sep="_")
feat_add <- feat_add %>%
  arrange(group, final_feat_group)

ggplot(feat_add, 
       aes(x=as.character(group), y=feature_group_name_proper, color=agg_level, group=final_feat_group), show.legend=F) + 
  geom_point() + 
  geom_path(arrow = arrow(ends="last", angle=20, length=unit(0.1, "inches"), type="closed")) +
  facet_grid(cancer_type~agg_level) + 
  coord_flip()+
  theme_light()+
  geom_text(aes(label=n_features_added), nudge_x=0.2, nudge_y=-0.15)+
  scale_color_jco()
```


Save results

```{r}
roc <- roc_auc_ %>%
  select("cancer_type", "agg_level", "min_test_auc", "sd_test_auc", "median_test_auc", "mean_test_auc",
         "train_test_diff", "ci_lower_roc_auc", "ci_upper_roc_auc") %>%
  setNames(c("Cancer type",	"Labeling type", "Min test ROC AUC",	"SD test ROC AUC",	"Median test ROC AUC",	
             "Mean test ROC AUC", "Train-test difference ROC AUC", "Lower bound of confidence interval for mean test ROC AUC",	
             "Upper bound of confidence interval for mean test ROC AUC"))

rec_3 <- recall_stats[recall_stats$quantile == 0.03, ] %>%
  select(c("cancer_type", "agg_level","mean_recall", "mean_lift_recall", "ci_lower_lift_recall",
           "ci_upper_lift_recall",  "mean_precision","mean_lift_precision", "ci_lower_lift_precision", 
           "ci_upper_lift_precision")) %>%
  setNames(c("Cancer type",		"Labeling type", "Mean recall",	"Mean lift of recall",
  "Lower bound of confidence interval for mean lift of recall",	"Upper bound of confidence interval for mean  lift of recall",
  "Mean precision",	"Mean lift of precision",	"Lower bound of confidence interval for mean lift of precision",
  "Upper bound of confidence interval for mean  lift of precision"))

rec_5 <- recall_stats[recall_stats$quantile == 0.05, ] %>%
  select(c("cancer_type", "agg_level","mean_recall", "mean_lift_recall", "ci_lower_lift_recall",
           "ci_upper_lift_recall",  "mean_precision","mean_lift_precision", "ci_lower_lift_precision", 
           "ci_upper_lift_precision")) %>%
  setNames(c("Cancer type",		"Labeling type", "Mean recall",	"Mean lift of recall",
  "Lower bound of confidence interval for mean lift of recall",	"Upper bound of confidence interval for mean  lift of recall",
  "Mean precision",	"Mean lift of precision",	"Lower bound of confidence interval for mean lift of precision",
  "Upper bound of confidence interval for mean  lift of precision"))

q3 <- roc %>%
  inner_join(
    rec_3, by=c("Cancer type","Labeling type")
  )
q5 <- roc %>%
  inner_join(
    rec_5, by=c("Cancer type", "Labeling type")
  )

# save

# wb <- createWorkbook()
# addWorksheet(wb, "q3")
# addWorksheet(wb, "q5")
# 
# writeData(wb, sheet="q3", q3)
# writeData(wb, sheet="q5", q5)

# saveWorkbook(wb, "../paper_materials/table15.xlsx", overwrite = T)

```

```{r}

# add base feature group column!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
fs_save <- fs %>%
  ungroup() %>%
  inner_join(
    df_f, by=c("feat_added"="feature")
  ) %>%
  select(cancer_type, agg_level, feature_selection_iter, feat_added, feature_group_name_proper,  
         mean_lift_recall_003, mean_te_auc, mean_tr_auc) %>%
  setNames(c("Cancer type",	"Labeling type", "Feature selection iteration", "Feature added", 
             "Feature added group",  "Mean lift of recall (0.03)", "Mean test ROC AUC",
             "Mean train ROC AUC"))

# save

# wb <- createWorkbook()
# addWorksheet(wb, "selected_features")
# writeData(wb, sheet="selected_features", fs_save)
# saveWorkbook(wb, "../paper_materials/table16.xlsx", overwrite = T)
```

