---
title: "Classification report"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r "loading data", echo=F}
library(ggplot2)
library(ggsci)
library(wesanderson)
library("ggdendro")

library(dplyr)
library(reshape2)

library(formattable)

library(openxlsx)


win_len <- 100000
data_path <- "../data/output/classifier_results_best_features/"
roc_auc_data <- read.csv(paste0(data_path, "result_roc_auc_100000.csv"), row.names = 1, stringsAsFactors = F)
recall_data <- read.csv(paste0(data_path, "result_recall_100000.csv"), row.names = 1, stringsAsFactors = F)
imp_data <- read.csv(paste0(data_path, "result_imp_100000.csv"), row.names = 1, stringsAsFactors = F)

```

### ROC AUC analysis

```{r}
all_features_data <- roc_auc_data


roc_auc_cancer_stats <- all_features_data %>%
  group_by(cancer_type) %>%
  summarize(
    min_test_auc = min(te_auc),
    sd_test_auc = sd(te_auc),
    test_auc = median(te_auc),
    test_auc_mean = mean(te_auc),
    n_auc = n(),
    train_auc = median(tr_auc)
  )

roc_auc_cancer_stats %>% 
  arrange(min_test_auc)
```


```{r fig.width=10}
ggplot(all_features_data, aes(te_auc)) +
  geom_density(alpha=0.5)+
  facet_wrap(~ cancer_type, nrow = 2)
```

Difference between train and test

```{r fig.width=10}
all_features_data$diff_auc <- all_features_data$tr_auc - all_features_data$te_auc
ggplot(all_features_data, aes(diff_auc)) +
  geom_density(alpha=0.5)+
  facet_wrap(~ cancer_type, nrow = 2) 
```

```{r echo=F}
by_hspt <- all_features_data %>%
  mutate(
    train_test_diff = tr_auc - te_auc
  ) %>%
  group_by(cancer_type, agg_level) %>%
  summarize(
    min_test_auc = min(te_auc),
    sd_test_auc = sd(te_auc),
    test_auc = median(te_auc),
    test_auc_mean = mean(te_auc),
    train_test_diff = mean(train_test_diff)
  ) 
```

### Quality of clasifiers analysis

Aggregated statistics for hotspots prediction by cancer type

Recall

```{r fig.width=10}
all_features_data <- recall_data

all_features_data$agg_level <- as.character(all_features_data$agg_level)
all_features_data$quantile_chr <- as.character(all_features_data$quantile)

q <- unique(all_features_data$quantile_chr)
```

One panel for different probability quantiles

0.01

```{r fig.width=10}
map_df <- data.frame(agg_level=c("99.", "99.5.", "99.9."), agg_level_num = c(0.01, 0.005, 0.001))
all_features_data <- all_features_data %>%
  inner_join(map_df)
all_features_data$lift_precision <- all_features_data$precision / all_features_data$agg_level_num
comp_rec <- all_features_data %>%
  filter(quantile_chr == "0.01") 

comp_rec_pl <- comp_rec %>%
  select('cancer_type', 'agg_level', 'lift_recall', 'lift_precision') %>%
  melt(id.vars=c("cancer_type","agg_level"))

ggplot(comp_rec_pl, aes(x=cancer_type, y=value, fill=agg_level)) + 
  geom_boxplot() + 
  facet_wrap(~variable, scales = "free", nrow = 4) + 
  scale_fill_jco()
```

0.03

```{r fig.width=10}
comp_rec <- all_features_data %>%
  filter(quantile_chr == "0.03") 

comp_rec_pl <- comp_rec %>%
  select('cancer_type', 'agg_level', 'lift_recall', 'lift_precision') %>%
  melt(id.vars=c("cancer_type","agg_level"))

ggplot(comp_rec_pl, aes(x=cancer_type, y=value, fill=agg_level)) + 
  geom_boxplot() + 
  facet_wrap(~variable, scales = "free", nrow = 4) + 
  scale_fill_jco()
```
Confidence intervals for lift of recall / recall for different quantiles

```{r}
ci_stats <- all_features_data %>%
  group_by(cancer_type, agg_level, quantile) %>%
  summarize(
    mean_recall = mean(recall),
    mean_lift_recall = mean(lift_recall),
    mean_precision = mean(precision),
    mean_lift_precision = mean(lift_precision),
    sd_recall = sd(recall),
    sd_lift_recall = sd(lift_recall),
    sd_precision = sd(precision),
    sd_lift_precision = sd(lift_precision),
    n_ex = n()
  ) %>%
  mutate(
    ci_lower_recall = mean_recall - 1.96 * sd_recall / sqrt(n_ex),
    ci_upper_recall = mean_recall + 1.96 * sd_recall / sqrt(n_ex),
    
    ci_lower_lift_recall = mean_lift_recall - 1.96 * sd_lift_recall / sqrt(n_ex),
    ci_upper_lift_recall = mean_lift_recall + 1.96 * sd_lift_recall / sqrt(n_ex),    
    
    ci_lower_precision = mean_precision - 1.96 * sd_precision / sqrt(n_ex),
    ci_upper_precision = mean_precision + 1.96 * sd_precision / sqrt(n_ex),       
    
    ci_lower_lift_precision = mean_lift_precision - 1.96 * sd_lift_precision / sqrt(n_ex),
    ci_upper_lift_precision = mean_lift_precision + 1.96 * sd_lift_precision / sqrt(n_ex)
  )

```


```{r save results}
# roc auc
roc_data <- roc_auc_data %>%
  mutate(
    train_test_diff = tr_auc - te_auc
  ) %>%
  group_by(cancer_type, agg_level) %>%
  summarize(
    min_test_auc = min(te_auc),
    sd_test_auc = sd(te_auc),
    median_test_auc = median(te_auc),
    mean_test_auc = mean(te_auc),
    train_test_diff = mean(train_test_diff),
    n_ex = n()
  ) %>%
  mutate(
    ci_lower_roc_auc = mean_test_auc - 1.96 * sd_test_auc / sqrt(n_ex),
    ci_upper_roc_auc = mean_test_auc + 1.96 * sd_test_auc / sqrt(n_ex), 
  ) %>%
  select(-n_ex)

# recall
recall_stats_hsp <- ci_stats %>% 
  filter(quantile %in% c(0.001, 0.005, 0.01,0.02,0.03, 0.040, 0.050, 0.100)) %>%
  select(cancer_type, agg_level, quantile, mean_recall, mean_lift_recall, mean_precision, mean_lift_precision, ci_lower_lift_recall, ci_upper_lift_recall, ci_lower_lift_precision,ci_upper_lift_precision)


# save
# wb <- createWorkbook()
# addWorksheet(wb, "ROC_AUC_hotspots")
# addWorksheet(wb, "Recall_hotspots")
# 
# writeData(wb, sheet="ROC_AUC_hotspots", roc_data)
# writeData(wb, sheet="Recall_hotspots", recall_stats_hsp)
# 
# saveWorkbook(wb, "../reports/train_best_features.xlsx", overwrite = T)
```

Compare with results on all features


Median ROC AUC

```{r}
roc_data_base <- read.xlsx("../reports/train_baseline.xlsx", sheet = "ROC_AUC_hotspots")
recall_data_base <- read.xlsx("../reports/train_baseline.xlsx", sheet = "Recall_hotspots")


roc_med <- roc_data_base %>% 
  select(cancer_type, agg_level, median_test_auc) %>%
  setNames(c("cancer_type", "agg_level", "median_test_auc_base")) %>%
  inner_join(
    roc_data %>%
      select(cancer_type, agg_level, median_test_auc) %>%
      setNames(c("cancer_type", "agg_level", "median_test_auc_reduced")), by=c('cancer_type', 'agg_level'))
  
roc_med  
```



```{r}
roc_med_pl <- melt(roc_med,id.vars = c("cancer_type", "agg_level"))
roc_med_pl$variable_new <- gsub(x = roc_med_pl$variable, pattern = "median_test_auc_", replacement = "")
ggplot(roc_med_pl, aes(x=cancer_type, y=value, fill=variable_new)) + 
  geom_bar(stat='identity',  color="black", position = "dodge") +
  scale_fill_jco()
```


Train test diff ROC AUC

```{r}
roc_data_base %>% 
  select(cancer_type, agg_level, train_test_diff) %>%
  setNames(c("cancer_type", "agg_level", "train_test_diff_base")) %>%
  inner_join(
    roc_data %>%
      select(cancer_type, agg_level, train_test_diff) %>%
      setNames(c("cancer_type", "agg_level", "train_test_diff_reduced")), by=c('cancer_type', 'agg_level'))
```

SD ROC AUC

```{r}
roc_data_base %>% 
  select(cancer_type, agg_level, sd_test_auc) %>%
  setNames(c("cancer_type", "agg_level", "sd_test_auc_base")) %>%
  inner_join(
    roc_data %>%
      select(cancer_type, agg_level, sd_test_auc) %>%
      setNames(c("cancer_type", "agg_level", "sd_test_auc_reduced")), by=c('cancer_type', 'agg_level'))
```

Comparison of quality of classifiers

```{r}
recall_all <- recall_data_base %>%
  select(cancer_type, agg_level, quantile, mean_recall,mean_lift_recall,
             mean_precision, mean_lift_precision) %>%
  setNames(c("cancer_type", "agg_level", "quantile", "mean_recall_base","mean_lift_recall_base",
             "mean_precision_base", "mean_lift_precision_base")) %>%
  inner_join(
    recall_stats_hsp %>%
      select(cancer_type, agg_level, quantile, mean_recall, mean_lift_recall,
             mean_precision, mean_lift_precision) %>%
        setNames(c("cancer_type", "agg_level", "quantile", "mean_recall_reduced","mean_lift_recall_reduced",
             "mean_precision_reduced", "mean_lift_precision_reduced")),
    by=c('cancer_type', 'agg_level', 'quantile'))
recall_all$quantile_chr <- as.character(recall_all$quantile)
```

For 0.01 probability quantile

```{r  fig.width=10}
comp_rec <- recall_all %>%
  filter(quantile_chr == "0.01") %>%
  select("cancer_type", "mean_lift_recall_base", "mean_lift_recall_reduced", "mean_lift_precision_base", "mean_lift_precision_reduced")

comp_rec_pl <- comp_rec %>%
  melt(id.vars=c("cancer_type"))
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "prec"), "variable_type"] <- "precision"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "recall"), "variable_type"] <- "recall"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "base"), "variable_new"] <- "base"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "reduced"), "variable_new"] <- "reduced"
ggplot(comp_rec_pl, aes(x=cancer_type, y=value, fill=variable_new)) + 
  geom_bar(stat='identity', color="black", position = "dodge") +
  facet_wrap(~variable_type,nrow = 2) + 
  scale_fill_jco()
```

For 0.03 probability quantile

```{r  fig.width=10}
comp_rec <- recall_all %>%
  filter(quantile_chr == "0.03") %>%
  select("cancer_type", "mean_lift_recall_base", "mean_lift_recall_reduced", "mean_lift_precision_base", "mean_lift_precision_reduced")

comp_rec_pl <- comp_rec %>%
  melt(id.vars=c("cancer_type"))
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "prec"), "variable_type"] <- "precision"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "recall"), "variable_type"] <- "recall"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "base"), "variable_new"] <- "base"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "reduced"), "variable_new"] <- "reduced"
ggplot(comp_rec_pl, aes(x=cancer_type, y=value, fill=variable_new)) + 
  geom_bar(stat='identity',  color="black", position = "dodge") +
  facet_wrap(~variable_type,nrow = 2) + 
  scale_fill_jco()
```

For 0.05 probability quantile

```{r  fig.width=10}
comp_rec <- recall_all %>%
  filter(quantile_chr == "0.05") %>%
  select("cancer_type", "mean_lift_recall_base", "mean_lift_recall_reduced", "mean_lift_precision_base", "mean_lift_precision_reduced")

comp_rec_pl <- comp_rec %>%
  melt(id.vars=c("cancer_type"))
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "prec"), "variable_type"] <- "precision"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "recall"), "variable_type"] <- "recall"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "base"), "variable_new"] <- "base"
comp_rec_pl[grep(x = comp_rec_pl$variable, pattern = "reduced"), "variable_new"] <- "reduced"
ggplot(comp_rec_pl, aes(x=cancer_type, y=value, fill=variable_new)) + 
  geom_bar(stat='identity',  color="black", position = "dodge") +
  facet_wrap(~variable_type,nrow = 2) + 
  scale_fill_jco()
```
